---
title: "Optimization: Exercises 4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Finite Differencing

In the slides, Forward-Difference (`df.FD`) was compared to exact differentiation (`df.exact`) for

Reconstruct (implement) the example and the results in the slides yourself.

```{r}
f = function(x) cos(sin(x)*cos(x))
df.exact = function(x) -cos(2*x)*sin(sin(x)*cos(x))
df.FD = function(x, h) (f(x + h) - f(x))/h
df.CD = function(x, h) (f(x+h) - f(x-h))/(2*h)
x <- pi/3
df_exact <- df.exact(x) ; df_exact
```

Include Central-Difference calculations in both trace and plots.
Remember to also add CD reference lines:

```{r}
eps <- .Machine$double.eps
cd_candidate <- fd_candidate <- h <- numeric(16)


for (i in 1:16){
  h[i] <- 10^(-i)
  fd_candidate[i] <- df.FD(x,h[i])
  cd_candidate[i] <- df.CD(x,h[i])}

# Plotter både CD = rød og FD = blå i samme plot.
plot(log10(h),log10(abs(cd_candidate - df_exact)) , type = "l" , col = "blue")
lines(log10(h),log10(abs(fd_candidate - df_exact)) , col = "red")

abline(v = log10((sqrt(eps))), col = "yellow")
abline(h = log10((sqrt(eps))), col = "yellow")
```

Calculate the order ratio for CD and compare to that of FD.

```{r}

```

Try out Richardson Extrapolation for CD and compare to that of FD.

```{r}

```

What happens if we misuse the Richardson Extrapolation:

* For (incorrect) first-order correction for CD?
* For (incorrect) second-order correction for FD?
* Investigate for different choices of $h$.

```{r}

```
